---
title: "BankN Case Study"
author: 
  - Nathaniel Bedassie
  - Jheanel Brown
  - Taffeya Dunkley
  - Phillip Llewelyn
  - Tai-vaughn White
date: "6/18/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 9999)
knitr::opts_chunk$set(linewidth = 60)
```

## Question 1
1. Load the dataset and explore the contents, summarizing your findings of the data quality in
relation to standard measures (including, mean, median, standard deviation, missing values,
outliers, etc.). Review the data dictionary attached to get the context of each measurement.

    a.Describe the data in terms of the types and measurements and what each measurement
represents (purpose of features, range of values, min, max, etc.)

    b.Fix missing values, noise and outliers as necessary
    
```{r echo=FALSE}
bankn <- read.csv("2021-Set03-QualityAssured.csv", 
                  header = T, sep = ",", stringsAsFactors = T)
```

## Summary of Data 

### RefNum
```{r}
summary(bankn$RefNum)
```


### Age
```{r}
summary(bankn$age)
```

## Summary of Data Cont'd

### Job
```{r}
summary(bankn$job)
```

### Maritial Status
```{r}
summary(bankn$marital)
```

## Summary of Data Cont'd

### Education 
```{r}
summary(bankn$education)
```


### Housing
```{r}
summary(bankn$housing)
```

## Summary of Data Cont'd

### Loan
```{r}
summary(bankn$loan)
```


### Contact
```{r}
summary(bankn$contact)
```

## Summary of Data Cont'd

### Day
```{r}
summary(bankn$day)
```


### Month
```{r} 
summary(bankn$month)
```

## Summary of Data Cont'd

### Duration
```{r}
summary(bankn$duration)
```
- Std. Deviation
`r sd(bankn$duration)`


### Deposit
```{r}
summary(bankn$deposit)
```

## Summary of Data Cont'd

### Balance
```{r}
summary(bankn$balance, maxsum = 10)
```
- The balance field contains 4 rows with characters instead of numeric 
  at row 8, 54, 26 and 87. Which causes the field to be interpreted as a factor field.
`r bankn$balance[c(8, 54, 26, 87)]`



```{r echo=FALSE}
str(bankn)
levels(bankn$job)
```










