---
title: "BankN Case Study"
author: 
  - Nathaniel Bedassie
  - Jheanel Brown
  - Taffeya Dunkley
  - Phillip Llewelyn
  - Tai-vaughn White
date: "6/18/2021"
output:
  ioslides_presentation:
    logo: QualityAssuredlogo.jpeg
autosize: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 9999)
options(width = 60)
knitr::opts_chunk$set(linewidth = 60)
library(knitr)
library(formatR)
library(ggplot2)
opts_chunk$set(tidy.opts=list(width.cutoff=40),tidy=TRUE)
mode <- function(v, empty = FALSE) {
  if(empty){
    v = v[v != ""]
  }
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

## Question 1
1. Load the dataset and explore the contents, summarizing your findings of the data quality in
relation to standard measures (including, mean, median, standard deviation, missing values,
outliers, etc.). Review the data dictionary attached to get the context of each measurement.

    a.Describe the data in terms of the types and measurements and what each measurement
represents (purpose of features, range of values, min, max, etc.)

    b.Fix missing values, noise and outliers as necessary
    
```{r echo=FALSE}
bankn <- read.csv("2021-Set03-QualityAssured.csv", 
                  header = T, sep = ",", stringsAsFactors = T)
```

## Summary of Data 

### RefNum
- ***A unique identifier for each customer***
```{r}
summary(bankn$RefNum)
```
As seen above the values range from 10023467 to 10024584, which accurately
reflects that there are 1118 records. These Reference numbers may be needed later on.
There does not appear to be any outliers or missing values.

## Summary of Data Cont'd

### Age
- ***numeric age of the customer***
```{r}
summary(bankn$age)
```
- Std. Deviation (returns NA because of missing values)

`r sd(bankn$age)`
- There appears to be no unusual ages present, however, there are missing values. 
We have chosen to replace them with the median age.

## Fixing the missing values in age
```{r}
bankn$age[is.na(bankn$age)] <- median(bankn$age, na.rm = T)
summary(bankn$age)
```

## Summary of Data Cont'd

### Maritial Status
- ***Marital status ("married","divorced","single"; note: "divorced" means divorced or widowed)***

```{r}
summary(bankn$marital)
```
Nothing appears to be extraordinary in the marital status field.

## Summary of Data Cont'd

### Education 
- ***("secondary","primary","tertiary")***
```{r}
summary(bankn$education)
```
For the education level, we have 94 empty values and 2 "?" characters that we
believe may represent an unknown value. To replace these values, we have 
decided it best to replace the missing values based on job. Certain types 
of jobs dominate certain education levels as seen here:.

## Education Distribution by Job

```{r, fig.cap="Fig.3 Primary education dist. by Jobs"}
ggplot(bankn[bankn$education == "tertiary",c(3,5)], aes(x="", y=job, fill=job)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +labs( title="Fig.1 Tertiary education dist. by Jobs")+ theme_void()
```

## Education Distribution by Job cont'd

```{r, fig.cap="Fig.3 Primary education dist. by Jobs"}
ggplot(bankn[bankn$education == "secondary",c(3,5)], aes(x="", y=job, fill=job)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +labs( title="Fig.2 Secondary education dist. by Jobs")+ theme_void()
```

## Education Distribution by Job cont'd

```{r, fig.cap="Fig.3 Primary education dist. by Jobs"}
ggplot(bankn[bankn$education == "primary",c(3,5)], aes(x="", y=job, fill=job)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +labs( title="Fig.3 Primary education dist. by Jobs")+ theme_void()
      
```

## Analysis
With this information we can assign an education level to match the prevalence 
of that job at that level.

## Helping Function
- This function allows us to get the most prevalent education level for a job

```{r}
geteducation <- function(jobs){
  lis <- vector()
  for (job in jobs) {
    tert <- length(bankn$education[bankn$education == "tertiary" & bankn$job == job])
    sec <- length(bankn$education[bankn$education == "secondary" & bankn$job == job])
    prim <- length(bankn$education[bankn$education == "primary" & bankn$job == job])
    len = nrow(bankn[bankn$job == job,])
    lis <- append(lis, ifelse(tert > sec & tert > prim, "tertiary", ifelse(sec > tert & sec > prim,
           "secondary", ifelse(prim > tert & prim > sec,
           "primary", '?'))))
  }
  return(lis)
}
```

## Applying the function to Fix Errors/Unknown


```{r}
missingEducation <- bankn[!(bankn$education %in% 
                    c('tertiary','secondary','primary')),
                  c('job','education')]

bankn$education[!(bankn$education %in% 
                    c('tertiary','secondary','primary'))] <- geteducation(missingEducation$job)
summary(bankn$education)
```

## Summary of Data Cont'd

### Job
- <font size="3">***Type of Job ("admin", "unemployed", "management","housemaid", "entrepreneur", "student", "blue-collar","self-employed", "retired", "technician", "services")***</font>

```{r}
summary(bankn$job)
```
- <font size="3">According to the data dictionary, there should be only 11 levels,
but there is one extra level. A " " character. There are 5 records with the job.
What may be done to correct it, is to match a possible job based on 
their deposits and balance compared to persons with jobs that have similar deposits
and balances.</font>


## Summary of Data Cont'd

### Housing
```{r}
summary(bankn$housing)
```
- The housing data is categorical, so no standard values were obtained from it, it simply gave a count of yes or no. There were also no outlier or missing values.

## Summary of Data Cont'd

### Loan
```{r}
summary(bankn$loan)
levels(bankn$loan)
```
- The loan data is representing a boolean if the user has a loan or not
- It is represented as categorical variable with values "yes" or "no"
- There are 15 missing values that are read as a blank field in the dataframe.
- This data column has good data as 1103 of the field data value is filled

### Contact
```{r}
summary(bankn$contact)
```
- The contact column is empty and has no value. It cannot be used in analysis.

## Summary of Data Cont'd

### Day
```{r}
summary(bankn$day)
```
- The standard values obtained from the date were not exactly meaningful, as they do represent ordinal values for the days in a month. However when day was analyzed with the month and potential year of the data set, it was observed that there may be outlier for the month of February where the day was 29. This is possible if the year of the data set was 2019, since that was not a leap year. A definite outlier would have been where the day was 30 for the month of February. 

### Month
```{r} 
summary(bankn$month)
```

## Summary of Data Cont'd

### Duration
```{r}
summary(bankn$duration)
```
- Std. Deviation
`r sd(bankn$duration)`


### Deposit
```{r}
summary(bankn$deposit)
```

## Summary of Data Cont'd

### Balance
```{r}
summary(bankn$balance, maxsum = 10)
```

- Std. Deviation won't work because of characters. Thats why the data type is factor and not numeric

`r #sd(bankn$balance)`
- The balance field contains 4 rows with characters instead of numeric 
  at row 8, 54, 26 and 87. Which causes the field to be interpreted as a factor field.

## Values  
```{r echo=TRUE}
bankn$balance[c(8, 54, 26, 87)]
```


## Structure of the data
```{r echo=FALSE}
str(bankn)
#levels(bankn$job)
```










