---
title: "BankN Case Study"
author: 
  - Nathaniel Bedassie
  - Jheanel Brown
  - Taffeya Dunkley
  - Phillip Llewelyn
  - Tai-vaughn White
date: "6/18/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 9999)
knitr::opts_chunk$set(linewidth = 60)
```

## Question 1
1. Load the dataset and explore the contents, summarizing your findings of the data quality in
relation to standard measures (including, mean, median, standard deviation, missing values,
outliers, etc.). Review the data dictionary attached to get the context of each measurement.

    a.Describe the data in terms of the types and measurements and what each measurement
represents (purpose of features, range of values, min, max, etc.)

    b.Fix missing values, noise and outliers as necessary
    
```{r echo=FALSE}
bankn <- read.csv("2021-Set03-QualityAssured.csv", 
                  header = T, sep = ",", stringsAsFactors = T)
```

## Summary of Data 

### RefNum
```{r}
summary(bankn$RefNum)
```


### Age
```{r}
summary(bankn$age)
```
- Std. Deviation
`r sd(bankn$age)`

## Summary of Data Cont'd

### Job
```{r}
summary(bankn$job)
```

### Maritial Status
```{r}
summary(bankn$marital)
```

## Summary of Data Cont'd

### Education 
```{r}
summary(bankn$education)
```


### Housing
```{r}
summary(bankn$housing)
```
- The housing data is categorical, so no standard values were obtained from it, it simply gave a count of yes or no. There were also no outlier or missing values.

## Summary of Data Cont'd

### Loan
```{r}
summary(bankn$loan)
```


### Contact
```{r}
summary(bankn$contact)
```


## Summary of Data Cont'd

### Day
```{r}
summary(bankn$day)
```
- The standard values obtained from the date were not exactly meaningful, as they do represent ordinal values for the days in a month. However when day was analyzed with the month and potential year of the data set, it was observed that there may be outlier for the month of February where the day was 29. This is possible if the year of the data set was 2019, since that was not a leap year. A definite outlier would have been where the day was 30 for the month of February. 

### Month
```{r} 
summary(bankn$month)
```

## Summary of Data Cont'd

### Duration
```{r}
summary(bankn$duration)
```
- Std. Deviation
`r sd(bankn$duration)`


### Deposit
```{r}
summary(bankn$deposit)
```

## Summary of Data Cont'd

### Balance
```{r}
summary(bankn$balance, maxsum = 10)
```

- Std. Deviation
`r #sd(bankn$balance)`

- The balance field contains 4 rows with characters instead of numeric 
  at row 8, 54, 26 and 87. Which causes the field to be interpreted as a factor field.
`r bankn$balance[c(8, 54, 26, 87)]`



```{r echo=FALSE}
str(bankn)
levels(bankn$job)
```










